Measurement and Probability in Fuzzy Quantum Theories

### 2017-02-13 prop_iop.tex
From: Tai, Yu-Tsung Sent: Monday, February 13, 2017 2:23 To: Hanson, Andrew J. <hansona@indiana.edu> Cc: Sabry, Amr A. <sabry@indiana.edu>; Ortiz, Gerardo <ortizg@indiana.edu> Subject: RE: Belief: RE: Need a definition.
Updated appendix C, and hope it would be more readable than the previous
one...

About the Dutch book, I haven't found a detailed computation in Prof.
Busemeyer papers, but I found some in quantum Bayesian's paper
(http://link.aps.org/doi/10.1103/RevModPhys.85.1693)... Reading the
details until section 2... For the part I read so far, it seems that
they consider the Born rule a

\begin{proof}When $\mathscr{I}=\set{\left\{ a\right\} }{a\in\left[0,1\right]}$,
$\bar{\mu}$ and $\bar{\mu}^{\Omega}$ are essentially quantum and
classical real-valued probability measures, respectively. Hence, it
is sufficient to prove the interval-valued case.
\begin{itemize}
\item $\bar{\mu}^{\Omega}(\emptyset)=\bar{\mu}\left(\sum_{\ket{j}\in\emptyset}\proj{j}\right)=\bar{\mu}\left(\mathbb{0}\right)=[0,0]$. 
\item $\bar{\mu}^{\Omega}(\Omega)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)=[1,1]$. 
\item For any event $E\in2^{\Omega}$,
\begin{equation}\eqalign{ 
\bar{\mu}^{\Omega}\left(\Omega\backslash E\right)&=\bar{\mu}\left(\sum_{\ket{j}\in\Omega\backslash E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}-\sum_{\ket{j}\in E}\proj{j}\right)\\
&=\bar{\mu}\left(\mathbb{1}\right)-\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)=\left[1,1\right]-\bar{\mu}^{\Omega}\left(E\right)\textrm{ .} 
}\end{equation}
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}\subseteq2^{\Omega}$
of pairwise disjoint events, we have
\begin{equation}\eqalign{ 
\bar{\mu}^{\Omega}\left(\bigcup_{i=1}^{N}E_{i}\right)&=\bar{\mu}\left(\sum_{\ket{j}\in\bigcup_{i=1}^{N}E_{i}}\proj{j}\right)=\bar{\mu}\left(\sum_{i=1}^{N}\sum_{\ket{j}\in E_{i}}\proj{j}\right)\\
&\subseteq\sum_{i=1}^{N}\bar{\mu}\left(\sum_{\ket{j}\in E_{i}}\proj{j}\right)=\sum_{i=1}^{N}\bar{\mu}\left(E_{i}\right)\textrm{ .} 
}\end{equation}
\end{itemize}
\end{proof}

\begin{definition}Given a Hilbert space $\Hilb$, $\events'\subseteq\events$
is called a sub-event if it satisfies the follow conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$.
\item $\mathbb{1}\in\events'$.
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$.
\end{itemize}
By replacing every occurrence of $\events$ by $\events'$ in definition~\ref{def:QuantumProbabilitySpace}
and \ref{def:QuantumInterval-valuedProbability}, it is straightforward
to define a quantum real-valued and interval-valued probability measure
on $\events'$.\end{definition}

\begin{thm}Given a Hilbert space~$\Hilb$ and a mutually commuting
sub-event~$\events'\subseteq\events$, there is set~$\Omega$ and
a function~$\tau:\events'\rightarrow2^{\Omega}$ such that:
\begin{itemize}
\item If $\mu:\events'\rightarrow\left[0,1\right]$ is a quantum real-valued
probability measure, then there is a classical real-valued probability
measure~$\mu':\Omega\rightarrow\left[0,1\right]$ such that $\mu\left(P\right)=\mu'\left(\tau\left(P\right)\right)$
for all $P\in\events'$.
\item If $\mu:\events'\rightarrow\mathscr{I}$ is a quantum interval-valued
probability measure, then there is a classical interval-valued probability
measure~$\mu':\Omega\rightarrow\mathscr{I}$ such that $\mu\left(P\right)=\mu'\left(\tau\left(P\right)\right)$
for all $P\in\events'$
\end{itemize}
\end{thm}

\begin{proof}Again, it is sufficient to prove the interval-valued
case. Since $\events'$ is a set of mutually orthogonal projections,
they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
Pick arbitrary projector~$P\in\events'$, its eigenvalues are $0$
or $1$. Let $\tau\left(P\right)\subseteq\Omega$ be the set of eigenvectors
whose eigenvalues are $1$. Then, we have $P=\sum_{\ket{j}\in\tau\left(P\right)}\proj{j}$.
By definition~\ref{definition:quantum-classical}, $\mu\left(P\right)=\mu^{\Omega}\left(\tau\left(P\right)\right)$,
and $\mu^{\Omega}$ is a quantum interval-valued probability measure
by lemma~\ref{lemma:classical-probability-measures}.\end{proof}

By the previous lemma, given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$,
if we only consider the events generated by $\Omega$, $\events^{\Omega}=\set{\sum_{\ket{j}\in E}\proj{j}}{E\subseteq\Omega}$,
then any quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
behaves classically when restricting the domain on $\events^{\Omega}$,
i.e., consider $\bar{\mu}|_{\events^{\Omega}}$. In particular, if
$\bar{\mu}$ is convex, each $\bar{\mu}|_{\events^{\Omega}}$ is convex
and has a non-empty core. However, the core elements of each $\bar{\mu}|_{\events^{\Omega}}$
may not be glued together, and gives a core element of $\bar{\mu}$.
The inability of gluing core elements among commuting $\events^{\Omega}$
to non-commuting $\events$ is consistent with the idea of the Kochen-Specker
theorem, that each triple of commuting observables could be colored
well, but we cannot color all observables which are not all commuting.
This is also consistent with the assertion that mutually commuting
observables could be measured simultaneously~\cite{Mermin_1993}.
In another word, if Alice and Bob in Bell's theorem can only pick
from a set of mutually commuting observables, they could never observe
any ``quantum,`` and we cannot eliminate the local hidden variable
theory in this setting.

\begin{lemma}Given a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
if $\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ are all
convex for any orthonormal basis~$\Omega$, then $\bar{\mu}$ is
convex.\end{lemma}
\begin{proof}
Pick two commuting events~$P_{0},P_{1}\in\events$. Since they are
commuting, they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
Since they are projectors, the eigenvalues of them are $0$ or $1$
so that we can express them as $P_{0}=\sum_{\ket{j}\in E_{0}}\proj{j}$
and $P_{1}=\sum_{\ket{j}\in E_{1}}\proj{j}$, where $E_{0},E_{1}\subseteq\Omega$.
Then, 
\begin{equation}\eqalign{
 & \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\\
 = & \bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\vee\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)+\bar{\mu}\left(\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\right)\\
 = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cup E_{1}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{0}\cap E_{1}}\proj{j}\right)\\
 = & \bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\\
 \subseteq & \bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\\
 = & \bar{\mu}\left(\sum_{\ket{j}\in E_{0}}\proj{j}\right)+\bar{\mu}\left(\sum_{\ket{j}\in E_{1}}\proj{j}\right)\\
 = & \bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)
}\end{equation}
\end{proof}
\begin{lemma}When the number of elements in $\Omega$ is less than
or equal to 3, every classical IVPM~$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$
is convex.\end{lemma}
\begin{proof}
Given two events~$E_{0},E_{1}\subseteq\Omega$, we want to verify
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}\label{eq:classical-convex}
\end{equation}
When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). When
$E_{0}\cup E_{1}=\Omega$, we have $E_{0}^{c}\cap E_{1}^{c}=\emptyset$,
where $E_{0}^{c}=\Omega\backslash E_{0}$ and $E_{1}^{c}=\Omega\backslash E_{1}$.
Thus, 
\begin{equation}\eqalign{
 & \bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 \Rightarrow & \left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\
 \Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cup E_{1}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cap E_{1}^{c}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{1}^{c}\right)\right)\\
 \Rightarrow & \bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}
}\end{equation}
When the number of elements in $\Omega$ is less than or equal to
3, either $E_{0}\cap E_{1}=\emptyset$ or $E_{0}\cup E_{1}=\Omega$
must happen so that $\bar{\mu}^{\Omega}$ is convex.
\end{proof}
\begin{thm}Given a Hilbert space~$\Hilb$ of dimension~$3$, every
quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is convex.\end{thm}
\begin{proof}
The direct consequence of the previous lemmas.
\end{proof}

>>>>

\begin{proof}When $\mathscr{I}=\set{\left\{ a\right\} }{a\in\left[0,1\right]}$,
$\bar{\mu}$ and $\bar{\mu}^{\Omega}$ are essentially quantum and
classical real-valued probability measures, respectively. Hence, it
is sufficient to prove the interval-valued case.
\begin{itemize}
\item $\bar{\mu}^{\Omega}(\emptyset)=\bar{\mu}\left(\sum_{\ket{j}\in\emptyset}\proj{j}\right)=\bar{\mu}\left(\mathbb{0}\right)=[0,0]$. 
\item $\bar{\mu}^{\Omega}(\Omega)=\bar{\mu}\left(\sum_{\ket{j}\in\Omega}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)=[1,1]$. 
\item For any event $E\in2^{\Omega}$,
\begin{equation}
\bar{\mu}^{\Omega}\left(\Omega\backslash E\right)=\bar{\mu}\left(\mathbb{1}-\sum_{\ket{j}\in E}\proj{j}\right)=\bar{\mu}\left(\mathbb{1}\right)-\bar{\mu}\left(\sum_{\ket{j}\in E}\proj{j}\right)=\left[1,1\right]-\bar{\mu}^{\Omega}\left(E\right)\textrm{ .}
\end{equation}
\item For a collection $\left\{ E_{i}\right\} _{i=1}^{N}\subseteq2^{\Omega}$
of pairwise disjoint events, we have
\begin{equation}
\bar{\mu}^{\Omega}\left(\bigcup_{i=1}^{N}E_{i}\right)=\bar{\mu}\left(\sum_{i=1}^{N}\sum_{\ket{j}\in E_{i}}\proj{j}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(\sum_{\ket{j}\in E_{i}}\proj{j}\right)=\sum_{i=1}^{N}\bar{\mu}^{\Omega}\left(E_{i}\right)\textrm{ .}
\end{equation}
\end{itemize}
\end{proof}

\begin{definition}Given a Hilbert space $\Hilb$, $\events'\subseteq\events$
is called a sub-event space if it satisfies the follow conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$.
\item $\mathbb{1}\in\events'$.
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$.
\end{itemize}
By replacing every occurrence of $\events$ by $\events'$ in definition~\ref{def:QuantumProbabilitySpace}
and \ref{def:QuantumInterval-valuedProbability}, it is straightforward
to define a quantum real-valued and interval-valued probability measure
on $\events'$.\end{definition}

\begin{thm}Given a Hilbert space~$\Hilb$ and a mutually commuting
sub-event space~$\events'\subseteq\events$, there is set~$\Omega$
and a function~$\tau:\events'\rightarrow2^{\Omega}$ such that:
\begin{itemize}
\item If $\mu:\events'\rightarrow\left[0,1\right]$ is a quantum real-valued
probability measure, then there is a classical real-valued probability
measure~$\mu':\Omega\rightarrow\left[0,1\right]$ such that $\mu\left(P\right)=\mu'\left(\tau\left(P\right)\right)$
for all $P\in\events'$.
\item If $\bar{\mu}:\events'\rightarrow\mathscr{I}$ is a quantum interval-valued
probability measure, then there is a classical interval-valued probability
measure~$\bar{\mu}':\Omega\rightarrow\mathscr{I}$ such that $\bar{\mu}\left(P\right)=\bar{\mu}'\left(\tau\left(P\right)\right)$
for all $P\in\events'$
\end{itemize}
\end{thm}

\begin{proof}Again, it is sufficient to prove the interval-valued
case. Since $\events'$ is a set of mutually orthogonal projections,
they can be diagonalized by a common orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$.
Pick arbitrary projector~$P\in\events'$, its eigenvalues are $0$
or $1$. Let $\tau\left(P\right)\subseteq\Omega$ be the set of eigenvectors
whose eigenvalues are $1$. Then, we have $P=\sum_{\ket{j}\in\tau\left(P\right)}\proj{j}$.
By definition~\ref{definition:quantum-classical}, $\bar{\mu}\left(P\right)=\bar{\mu}^{\Omega}\left(\tau\left(P\right)\right)$,
and $\bar{\mu}^{\Omega}$ is a quantum interval-valued probability
measure by lemma~\ref{lemma:classical-probability-measures}.\end{proof}

Given an orthonormal basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$,
if we consider sub-event space generated by $\Omega$, $\events^{\Omega}=\set{\sum_{\ket{j}\in E}\proj{j}}{E\subseteq\Omega}$,
then any quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
restricting on $\events^{\Omega}$, denoted by $\bar{\mu}|_{\events^{\Omega}}$,
behaves classically. In particular, if $\bar{\mu}$ is convex, each
$\bar{\mu}|_{\events^{\Omega}}$ is convex and has a non-empty core.
However, the core elements of each $\bar{\mu}|_{\events^{\Omega}}$
may not be glued together, and gives a core element of $\bar{\mu}$.
The inability of gluing core elements among commuting $\events^{\Omega}$
to non-commuting $\events$ is consistent with the idea of the Kochen-Specker
theorem, that each triple of commuting observables could be colored
well, but we cannot color all observables which are not all commuting.
This is also consistent with the assertion that mutually commuting
observables could be measured simultaneously~\cite{Mermin_1993}.
In another word, if Alice and Bob in Bell's theorem can only pick
from a set of mutually commuting observables, they could never observe
any ``quantum,`` and we cannot eliminate the local hidden variable
theory in this setting.

The rest of the section focus on convexity.

\begin{lemma}Given a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$,
if $\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$ are all
convex for any orthonormal basis~$\Omega$, then $\bar{\mu}$ is
convex.\end{lemma}

\begin{proof} Pick two commuting events~$P_{0},P_{1}\in\events$.
Since they are commuting, they can be diagonalized by a common orthonormal
basis~$\Omega=\left\{ \ket{j}\right\} _{j=0}^{d-1}$. Since they
are projectors, the eigenvalues of them are $0$ or $1$ so that we
can express them as $P_{0}=\sum_{\ket{j}\in E_{0}}\proj{j}$ and $P_{1}=\sum_{\ket{j}\in E_{1}}\proj{j}$,
where $E_{0},E_{1}\subseteq\Omega$. Then, 
\begin{equation}\eqalign{  
& \bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)= \bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\\ 
\subseteq & \bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)= \bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)\textrm{ .} 
}\end{equation}
\end{proof}

\begin{lemma}When the number of elements in $\Omega$ is less than
or equal to 3, every classical IVPM~$\bar{\mu}^{\Omega}:2^{\Omega}\rightarrow\mathscr{I}$
is convex.\end{lemma}

\begin{proof}Given two events~$E_{0},E_{1}\subseteq\Omega$, we
want to verify 
\begin{equation}
\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .}\label{eq:classical-convex}
\end{equation}
When $E_{0}\cap E_{1}=\emptyset$, equation~(\ref{eq:classical-convex})
is a special case of equation~(\ref{eq:classical-include}). When
$E_{0}\cup E_{1}=\Omega$, we have $E_{0}^{c}\cap E_{1}^{c}=\emptyset$,
where $E_{0}^{c}=\Omega\backslash E_{0}$ and $E_{1}^{c}=\Omega\backslash E_{1}$.
Thus, 
\begin{equation}\eqalign{ 
& \bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\ 
\Rightarrow & \left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cup E_{1}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\cap E_{1}^{c}\right)\subseteq\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{0}^{c}\right)+\left[1,1\right]-\bar{\mu}^{\Omega}\left(E_{1}^{c}\right)\\ 
\Rightarrow & \bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cup E_{1}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\cap E_{1}^{c}\right)\right)\subseteq\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{0}^{c}\right)\right)+\bar{\mu}^{\Omega}\left(\Omega\backslash\left(E_{1}^{c}\right)\right)\\ 
\Rightarrow & \bar{\mu}^{\Omega}\left(E_{0}\cap E_{1}\right)+\bar{\mu}^{\Omega}\left(E_{0}\cup E_{1}\right)\subseteq\bar{\mu}^{\Omega}\left(E_{0}\right)+\bar{\mu}^{\Omega}\left(E_{1}\right)\textrm{ .} 
}\end{equation}
When the number of elements in $\Omega\le3$, either $E_{0}\cap E_{1}=\emptyset$
or $E_{0}\cup E_{1}=\Omega$ must happen so that $\bar{\mu}^{\Omega}$
is convex. \end{proof}

\begin{thm}Given a Hilbert space~$\Hilb$ of dimension~$3$, every
quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$ is convex.\end{thm}

\begin{proof}The direct consequence of the previous lemmas. \end{proof}
